---
title: "Intrahost model"
author: "JT McCrone"
date: "4/18/2017"
output: github_document
---

```{r,echo=F}
# Set up packages
require(knitr)
require(ggplot2)
require(magrittr)
require(tidyverse)
if(!("package:tidyverse" %in% search())){
  require(dplyr)
  require(tidyr)
  require(purrr)
  require(readr)
  require(rlang)
}
require(HIVEr)
require(extrafont)
require(wesanderson)
require(grid)
require(ggbeeswarm)
set.seed(42) # Set seed so randomization is reproducible

# Set up figures
opts_chunk$set(fig.align="center",warning=TRUE,tidy=T,cache = T,echo=F)
theme_set(new = theme_classic()+ theme(
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour ='black',size=0.5,linetype='solid'),
text=element_text(family="Arial",size = 18))) # to make nice plots
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbPalette<-wes_palette("Zissou")
# Get widely used functions
#source("../scripts/useful_functions.R")
require(doMC)
doMC::registerDoMC(cores=4)

####### Write to summary results file ######
write_to_summary<-function(line_pattern,value){
  file = readLines("./results.table.tsv")
  line_pattern_regex = paste0("^",line_pattern)
  line = grep(line_pattern_regex,file)
  file[line] = paste0(line_pattern,"\t",value)
  writeLines(file,"./results.table.tsv")
}
require(cowplot)
```

```{r}

meta<-read_csv("../data/reference/all_meta.sequence_success.csv")
intra<-read_csv("../data/processed/secondary/Intrahost_initially_present.csv")


intra %>% mutate(DPS1 = collect1-onset,DPS2 = collect2-onset) ->intra
intra<-subset(intra,freq1<0.5) # Just to make sure


intra %>% group_by(ENROLLID) %>% 
  summarize(onset = unique(onset),
            within_host_time = unique(within_host_time),
            DPS1 = unique(DPS1),
            DPS2 = unique(DPS2)) -> intra_meta

intra_meta<-intra_meta[order(intra_meta$DPS1,decreasing = T),]
intra_meta %>% mutate(DPS2 = ifelse(DPS1==DPS2,yes = DPS2+0.3,no = DPS2)) ->intra_meta
intra_meta$sort_order<-1:nrow(intra_meta)
fig_2A<-ggplot(intra_meta,aes(x = DPS1,xend=DPS2,y = sort_order,yend=sort_order))+geom_segment(color = cbPalette[1])+ylab("")+xlab("Day post symptom onset")+ theme(axis.line.y=element_blank(),axis.ticks.y = element_blank(),axis.text.y = element_blank())+geom_point(aes(y=sort_order,x = DPS1),color=cbPalette[1])+geom_point(aes(y = sort_order,x = DPS2),color=cbPalette[1]) + scale_x_continuous(breaks = -2:6) 

fig_2A

save_plot("./Figures/Figure2A.pdf", fig_2A,
          base_aspect_ratio = 1.3)
embed_fonts("./Figures/Figure2A.pdf")

write.csv(rename(intra_meta,day.post.sympotom.onset1=DPS1,
                 day.post.sympotom.onset2=DPS2,),"./Figures/data/Figure2A.csv")
```

```{r}
same_day<-subset(intra,within_host_time==0 & freq1<0.5)
Sup_6<-ggplot(same_day,aes(x=freq1,y=freq2))+geom_point()+
  xlab("Frequency in home isolate") + ylab("Frequency in clinic isolate") + 
  scale_x_continuous(limits = c(0,0.5))+scale_y_continuous(limits = c(0,0.5))
Sup_6

lm_fit<-lm(freq2~freq1,same_day)
summary(lm_fit)->sum_fit
sum_fit

write_to_summary("R2 samples same day:",sum_fit$r.squared)
save_plot("./Figures/Supplemental_Figure6.pdf", Sup_6,
          base_aspect_ratio = 1.3)
embed_fonts("./Figures/Supplemental_Figure6.pdf")

write.csv(select(same_day,SPECID1,SPECID2,collect1,collect2,chr,pos,ref,var,freq1,freq2),
          "./Figures/data/Supplemental_Figure6.csv")
```


## Within host dynamics
```{r}



stat_sum_df <- function(fun, geom="crossbar", ...) {
  stat_summary(fun.data = fun, geom = geom, width = 0.2, ...)
}

iqr_min <-function(x) quantile(x,probs = 0.25)
iqr_max <-function(x) quantile(x,probs = 0.75)

intra.plot<-ggplot(intra,aes(x=as.factor(within_host_time),y=freq2-freq1,fill=donor_class))+xlab("Time within host (days)")+ylab("Change in frequency")+
  scale_fill_manual(labels  =c( "Nonsynonymous","Synonymous"),
                    name="",values=cbPalette[c(1,4)])+
  geom_dotplot(stackdir = "center",binaxis = 'y',binwidth = 0.02,dotsize = 0.6,
               position=position_dodge(width = 0.95))+
  #stat_summary(fun.y = "median",fun.ymin =  "iqr_min",fun.ymax = "iqr_max",  geom = "crossbar",
  #             position = position_dodge(width = 1,preserve = 'single'),
  #             aes(color = donor_class,fill=NA))+
  #scale_color_manual(labels  ="",name="",values=cbPalette[c(1,4)])+
  theme(legend.position = 'none')

  
intra.plot
save_plot("./Figures/Figure2B.pdf", intra.plot,
          base_aspect_ratio = 1.3)

embed_fonts("./Figures/Figure2B.pdf")
write.csv(select(intra,days.betweeen.sampling=within_host_time,SPECID1,SPECID2,chr,pos,ref,var,
                 freq1,freq2,mutational.class=donor_class),"./Figures/data/Figure2B.csv")

intra %>% subset(within_host_time==1) %>% .$ENROLLID %>% unique() %>% length()->one_day_isnv
intra %>% subset(within_host_time>0) %>% .$ENROLLID %>% unique() %>% length()->at_least_day_isnv


write_to_summary("Samples exactly 1 day apart:",one_day_isnv)
write_to_summary("Samples at least 1 day apart:",at_least_day_isnv)
```


# Within host classes

```{r,fig.width=5}
intra_all<-read_csv("../data/processed/secondary/Intrahost_all.csv",col_types = cols(ENROLLID=col_character()))
# Remove mixed infections.
intra_all<- intra_all %>% filter(!(SPECID2 %in% c("HS1530","MH8137","MH8390")) & !(SPECID1 %in% c("HS1530","MH8137","MH8390")))

intra_all %>% mutate(DPS1 = collect1-onset,DPS2 = collect2-onset) ->intra_all
intra_all<-filter(intra_all,freq1<0.5) # Just to make sure

intra_all<-intra_all %>% mutate(Endpoint="Persistent") %>%
  mutate(Endpoint = if_else(freq1==0,"Arisen",Endpoint)) %>%
  mutate(Endpoint = if_else(freq2==0,"Lost",Endpoint))

intra_all$Endpoint<-factor(intra_all$Endpoint,levels = c("Persistent","Arisen","Lost"),ordered = T)

intra.plot<-ggplot(intra_all,aes(x=as.factor(within_host_time),
                             y=freq2-freq1,
                   fill = Endpoint))+
  xlab("Time within host (days)")+ylab("Change in frequency")+
  geom_quasirandom(pch=21,color='black',size=2)+scale_fill_manual(values=cbPalette[c(1,3,5)],name="")+facet_wrap(~class)
intra.plot

pdf("./Figures/Figure2B_2.pdf",width = 10,height = 7)
intra.plot
dev.off()



# save_plot("./Figures/Figure2B_2.pdf", intra.plot,
# #          base_aspect_ratio = 1.3,
#           base_width = )

embed_fonts("./Figures/Figure2B_2.pdf")
```


## Antigenic

```{r}
HA_nonsyn<-read_csv("../data/processed/secondary/minor_nonsynom.csv",col_types = cols(ENROLLID=col_character()))
HA_nonsyn<- HA_nonsyn %>% select(mutation,ENROLLID,mutation,H3_name,Antigenic,H1_name)

intra_all<-left_join(intra_all,HA_nonsyn)  %>% distinct()# add HA classification

HA_NS<-filter(intra_all,chr=="HA",(!is.na(H3_name)|!(is.na(H1_name))))
HA_NS <-mutate(HA_NS, class = ifelse(is.na(Antigenic),
                                     "Nonantigenic",
                                     "Antigenic"))
HA_NS.p<-ggplot(HA_NS,aes(x=as.factor(within_host_time),
                             y=freq2-freq1,
                   fill = Endpoint))+
  xlab("Time within host (days)")+ylab("Change in frequency")+
  geom_point(pch=21,color='black',size=2)+scale_fill_manual(values=cbPalette[c(1,3,5)],name="")+facet_wrap(~class)
HA_NS.p

```


# Comparing within to between dynamics

```{r}
intra<-read_csv("../data/processed/secondary/Intrahost_all.csv")
intra %>% mutate(DPS1 = collect1-onset,DPS2 = collect2-onset) ->intra
intra<-subset(intra,freq1<0.5) # Just to make sure



trans<-read_csv("../data/processed/secondary/trans_freq.csv")
trans<-trans %>% filter(freq1<0.5) %>%
  mutate(Endpoint="Persistent") %>%
  mutate(Endpoint = if_else(freq1==0,"Arisen",Endpoint)) %>%
  mutate(Endpoint = if_else(freq2==0,"Lost",Endpoint))
trans$Endpoint<-factor(trans$Endpoint,levels = c("Persistent","Arisen","Lost"),ordered = T)

trans <- trans %>% mutate(within_host_time = abs(transmission-collect1)+abs(collect2-transmission))

intra<-intra %>% mutate(Endpoint="Persistent") %>%
  mutate(Endpoint = if_else(freq1==0,"Arisen",Endpoint)) %>%
  mutate(Endpoint = if_else(freq2==0,"Lost",Endpoint))

intra$Endpoint<-factor(intra$Endpoint,levels = c("Persistent","Arisen","Lost"),ordered = T)

intra.plot<-ggplot(intra,aes(x=as.factor(within_host_time),
                             y=freq2-freq1,
                   color = Endpoint))+
  xlab("Time within host (days)")+ylab("Change in frequency")+
  geom_quasirandom()

trans.plot<-ggplot(trans,aes(x=as.factor(within_host_time),
                             y=freq2-freq1,
                   color = Endpoint))+
  xlab("Time within host (days)")+ylab("Change in frequency")+
  geom_quasirandom()

cowplot::plot_grid(intra.plot, trans.plot, labels = c("A", "B"), ncol = 2,align = c("v","h"))#+draw_label("DRAFT!", angle = 45, size = 80, alpha = .2)

ggplot(intra,aes(x=freq1,y=freq2))+geom_point(color=cbPalette[1])+geom_point(data=filter(trans,freq1<0.5),color=cbPalette[5])+geom_abline(slope = 1,intercept = 0,linetype=2,alpha=0.5)

```



```{r}
proportion<-intra %>% group_by(within_host_time) %>%
  summarize(Persistent =
              length(which(Endpoint=="Persistent"))/length(Endpoint),
            Arisen = 
              length(which(Endpoint=="Arisen"))/length(Endpoint),
            Lost = 
              length(which(Endpoint=="Lost"))/length(Endpoint))

ggplot(proportion,aes(x=within_host_time,y=Persistent))+geom_line()
```


### Are the iSNV independent

```{r,eval=F}
intra %>%  mutate(delta = abs(freq2-freq1))->intra

var_data<-data.frame(delta=intra$delta,freq2=intra$freq2,group = factor(intra$ENROLLID),time = factor(intra$within_host_time))

diffs = c()
group = c()
df = data.frame(diffs,group)
for(i in 1:(nrow(intra)-1)){
  for (j in (i+1):nrow(intra)){
    if(intra$within_host_time[i] == intra$within_host_time[j]){
      d=abs(intra$delta[i]-intra$delta[j])
      g = ifelse(test = intra$ENROLLID[i]==intra$ENROLLID[j],
                     yes = "Within_host",no = "Between_host")
      df<-rbind(df,data.frame(diffs=d,group=g))
    }
  }
}

ggplot(df,aes(x = diffs,y=..ncount..,fill=group))+
  geom_histogram(color="white",binwidth = 0.01,position = position_dodge()) +
  scale_fill_manual(labels  =c("Between hosts","Within host"),name="",values=cbPalette[c(1,4)])
  

ggplot(df,aes(y = diffs,x=group,fill=group))+
  geom_boxplot() +
  scale_fill_manual(labels  =c("Between hosts","Within host"),name="",values=cbPalette[c(1,4)])
wilcox.test(diffs~group,df,alternative='g')
```




If we pick 1/ person

```{r}
one_per<-read_csv("../data/processed/secondary/one_per_person.csv")
ggplot(one_per,aes(x= Ne))+geom_histogram(color="white",binwidth = 1)
summary(one_per)
```
### Simulations

```{r}
sim.df<-read_csv("./simulated_fits.csv")

write_to_summary("Simulations mean 30:",mean(sim.df$X30))
write_to_summary("Simulations mean 50:",mean(sim.df$X50))
write_to_summary("Simulations mean 100:",mean(sim.df$X100))

sim.df.l<-data.frame(outcomes = c(sim.df$X30,sim.df$X50,sim.df$X100),Ne = rep(c(30,50,100),each = nrow(sim.df)))

fig_2C<-ggplot(sim.df.l,aes(y = outcomes,x = as.factor(Ne)))+geom_boxplot()+scale_y_continuous(limits = c(0,240),breaks=seq(0,240,20))+xlab("Expected Ne")+ ylab("Estimated Ne") #geom_abline(slope=1,intercept = 0,linetype=2)+
fig_2C
save_plot("./Figures/Figure2C.pdf", fig_2C,
          base_aspect_ratio = 1.3)
embed_fonts("./Figures/Figure2C.pdf")
write.csv(sim.df.l,"./Figures/data/Figure2C.csv")


```

## Is the model appropriate?
What is the expected cumulative distribution of change in frequency?

First what is the cumulative distribution of the change in frequency for the data?
```{r}
fits<- intra %>% filter(within_host_time>0) %>% mutate(dfreq=abs(freq1-freq2))

cdf<-function(x){
  Fn<-ecdf(x)
  return(Fn(x))
}


fits<- fits %>% 
  mutate(cdf = cdf(dfreq))
ggplot(fits,aes(x=dfreq,y=cumulative))+geom_step()#+scale_x_continuous(limits=c(0,1))

#intra %>% filter(within_host_time<4) %>% ggplot(aes(x=dfreq,y=cdf,color=as.factor(within_host_time)))+geom_step()

```


# Expectations from simulations

```{r}
paste0("./sims_30/",list.files("./sims_30/"))->s30_files

s30<-read_rbind(s30_files,n=3)
s30<- s30 %>% mutate(dfreq=abs(freq1-sim))
s30<- s30 %>% 
  mutate(cdf = cdf(dfreq),Ne=30)

paste0("./sims_50/",list.files("./sims_50/"))->s50_files
s50<-read_rbind(s50_files,n=3)
s50<- s50 %>% mutate(dfreq=abs(freq1-sim))
s50<- s50 %>% 
  mutate(cdf = cdf(dfreq),Ne=50)


paste0("./sims_100/",list.files("./sims_100/"))->s100_files

s100<-read_rbind(s100_files,n=3)
s100<- s100 %>% mutate(dfreq=abs(freq1-sim))
s100<- s100 %>% 
  mutate(cdf = cdf(dfreq),Ne=100)
sims<-rbind(s30,s100,s50)
ggplot()+
  geom_step(data=sims,aes(x=dfreq,y=cdf,group=paste(run,Ne),color=as.factor(Ne)),alpha=0.3) + geom_step(data=fits,aes(x=dfreq,y=cdf)) +
  scale_x_log10()
#geom_step(data=s100,aes(x=dfreq,y=cdf,color="blue"))
```


###  Effect of removing data

```{r}
# copied from notebook 
#steps<-data.frame(Ne = c(32.0,37.0,42.0,50.0,52.0,54.0,56.0,59.0,60.0,61.0,67.0,70.0,76.0,77.0,86.0,109.0,135.0,140.0,154.0,155.0,150.0,151.0,154.0,165.0,175.0,183.0,190.0,197.0,188.0,195.0,201.0,191.0,181.0,202.0,207.0,241.0,226.0,215.0,253.0,239.0,222.0,273.0,453.0))
every_nth <- function(x, nth, empty = TRUE, inverse = FALSE) # Fromhttps://stackoverflow.com/questions/34533472/insert-blanks-into-a-vector-for-e-g-minor-tick-labels-in-r
  {
  if (!inverse) {
    if(empty) {
      x[1:nth == 1] <- ""
      x
      } else {
        x[1:nth != 1]
        }
    } else {
      if(empty) {
        x[1:nth != 1] <- ""
        x
        } else {
          x[1:nth == 1]
        }
    }
}

custom_breaks = seq(0,500,50)
#steps$removed<-0:(nrow(steps)-1)/(nrow(subset(intra,freq1<0.5))-nrow(same_day))
steps<-read.csv("../data/processed/secondary/removed_data.csv")
fig_2D<-ggplot(steps,aes(x=removed,y = Ne))+geom_line()+
  xlab("Fraction iSNV removed") + ylab("Estimated Ne")+
  geom_point(data=steps[1,])+
  #geom_abline(slope=0,intercept = steps$Ne[1]*10,linetype=2)+
  scale_x_continuous(limits = c(0,0.9),breaks = seq(0,0.9,0.1))+
  scale_y_continuous(limits = c(0,500),breaks = custom_breaks,
                     labels= every_nth(custom_breaks,2,inverse = T))
fig_2D
save_plot("./Figures/Figure2D.pdf", fig_2D,
          base_aspect_ratio = 1.3)
embed_fonts("./Figures/Figure2D.pdf")
write.csv(rename(steps,fraction.removed=removed),"./Figures/data/Figure2D.csv")

steps %>% filter(Ne>320)  ->ten_fold
first_above<- subset(ten_fold,Ne==min(ten_fold$Ne))
write_to_summary("Fraction to increase by 10-fold:",first_above$removed)
```
# Gif example

```{r,eval=F}
diff_ex.df<-read.csv("./diffusion_example.csv")

pdf(file="./Figures/diffusion_examples.pdf", width=10, height=10)
for (i in unique(diff_ex.df$Days)){
df = subset(diff_ex.df,Days==i)
df$Probability[df$Frequency %in% c(0,1)]=df$Probability[df$Frequency %in% c(0,1)]*20
print(ggplot(df,aes(x=Frequency,y=Probability))+geom_line(color=cbPalette[1],size=3)+scale_y_continuous(limits=c(0,5),labels = c(rep("",6)))+ylab("")+theme(axis.ticks.y = element_blank(),axis.title.x = element_text(size=36),axis.text.x = element_text(size=28) ))
}
dev.off()

system("convert -delay 20 ./Figures/diffusion_examples.pdf ./Figures/diffusion_examples.gif")
```
