---
title: "Transmission set up"
author: "JT McCrone"
date: "4/4/2017"
output: github_document
---

```{r,echo=F}

# Set up packages
require(knitr)
require(ggplot2)
require(magrittr)
require(tidyverse)
if(!("package:tidyverse" %in% search())){
  require(dplyr)
  require(tidyr)
  require(purrr)
  require(readr)
  require(rlang)
}
require(HIVEr)
require(extrafont)
require(wesanderson)
require(grid)
require(cowplot)
set.seed(42) # Set seed so randomization is reproducible

# Set up figures
opts_chunk$set(fig.align="center",warning=TRUE,tidy=T,cache = T,echo=F)
theme_set(new = theme_classic()+ theme(
axis.line.x = element_line(colour = 'black', size=0.5, linetype='solid'),
axis.line.y = element_line(colour ='black',size=0.5,linetype='solid'),
text=element_text(family="Arial",size = 18))) # to make nice plots
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbPalette<-wes_palette("Zissou")
# Get widely used functions
#source("../scripts/useful_functions.R")
require(doMC)
doMC::registerDoMC(cores=4)

####### Write to summary results file ######
write_to_summary<-function(line_pattern,value){
  file = readLines("./results.table.tsv")
  line_pattern_regex = paste0("^",line_pattern)
  line = grep(line_pattern_regex,file)
  file[line] = paste0(line_pattern,"\t",value)
  writeLines(file,"./results.table.tsv")
}
```


# Set up

Reading in the transmission pairs and the quality snv
```{r}
# Read in the meta data
meta<-read_csv("../data/reference/all_meta.sequence_success.csv")

qual<-read_csv("../data/processed/secondary/qual.snv.csv",
               col_types = list(
                 ENROLLID= col_character(),
                 SPECID = col_character(),
                 LAURING_ID = col_character(),
                 Id = col_character()
               ))
# Read in the transmission pairs made in the summary file
trans_pairs<-read_csv("../data/processed/secondary/transmission_pairs.csv")
trans_freq <- read_csv("../data/processed/secondary/trans_freq.csv")
```

```{r}
HA_nonsyn<-read_csv("../data/processed/secondary/minor_nonsynom.csv",col_types = cols(ENROLLID=col_character()))
HA_nonsyn<- HA_nonsyn %>% select(mutation,ENROLLID,mutation,H3_name,Antigenic,H1_name)

trans_freq<-left_join(trans_freq,HA_nonsyn)  %>% distinct()# add HA classification

trans_freq.p<-ggplot(trans_freq,aes(x=freq1,y=freq2))+geom_point()+xlab("Frequency in donor")+ylab("Frequency in recipient")+geom_point(data=filter(trans_freq,!is.na(Antigenic)),aes(color=cbPalette[5]))+theme(legend.position = "none")
trans_freq.p

# save_plot("./Figures/Figure3C.pdf", trans_freq.p,
#           base_aspect_ratio = 1.3)
# embed_fonts("./Figures/Figure3C.pdf")

write.csv(x = select(trans_freq,SPECID1,SPECID2,freq1,freq2),"./Figures/data/Figure3C.csv")
```
It is pecular that there are fixed differences between the donor and recipient. My hunch is that they are near the ends of the segment.



```{r}
recip_no_donor<-subset(trans_freq,freq1==0 & freq2>0.98)

print(nrow(recip_no_donor))
recip_no_donor$chr<-factor(recip_no_donor$chr,levels = rev(c("PB2","PB1","PA","HA","NP","NR","M","NS"))) # Set the segments as factors with PB2 on top
chrs<-read.csv("../data/reference/segs.csv",stringsAsFactors = F) # get the start and stop of each OR for each segment (2014-2015 used as reference)

chrs$chr<-factor(chrs$chr,levels=levels(recip_no_donor$chr)) # set factors on the is meta data

genome_loc.p<-ggplot(recip_no_donor,aes(x=pos,y=as.factor(chr)))+geom_point(shape=108,size=5,alpha=0.5)+geom_segment(data=chrs,aes(x = start, y = chr, xend = stop, yend = chr))+ylab("")+xlab("")+scale_color_manual(name="",values=cbPalette[c(1)])+ theme(axis.ticks =element_blank(),axis.line.x = element_blank(),axis.line.y=element_blank())+scale_x_continuous(breaks=c())+theme(legend.position = "none")
genome_loc.p+ggtitle("Fixed in Donor - lost in recipient")
```

Do these check out or do we think they are bugs.

Here I read in the raw variant calls.There are a lot so it takes some time.
```{r,eval=F}
variants_csv<-c("../data/processed/HK_1/all.variants.csv","../data/processed/HK_2/all.variants.csv","../data/processed/HK_6/all.variants.csv","../data/processed/HK_7/all.variants.csv","../data/processed/HK_8/all.variants.csv","../data/processed/cali09/all.variants.csv","../data/processed/cali09_2/all.variants.csv","../data/processed/victoria/all.variants.csv","../data/processed/victoria_2/all.variants.csv","../data/processed/perth/all.variants.csv","../data/processed/perth_2/all.variants.csv")
#
variants<-read_rbind(variants_csv)
#
# Id refers to the sequenced sample - LAURING_ID refers to the SPECID - the actual nose and through. There can be mulitple Id for the same LAURING_ID

LAURING_ID_LOOKUP<-data.frame(Id=unique(variants$Id))
LAURING_ID_LOOKUP<-adply(LAURING_ID_LOOKUP,1,function(x){
          LAURING_ID=strsplit(as.character(x$Id),"_")[[1]][1]
          dup = strsplit(as.character(x$Id),"_")[[1]][2] # get the duplicate label if needed otherwise returns NA
          
          if(is.na(as.numeric(LAURING_ID))==F){ # some Ids include a decimal - this removes that
            LAURING_ID=as.character(round(as.numeric(LAURING_ID),0))
          }
          
          x$LAURING_ID=LAURING_ID
          x$dup = dup
          return(x)})


variants<-join(variants,LAURING_ID_LOOKUP)
variants<-subset(variants,select = -c(X,Unnamed..0.1))
variants<-join(variants,meta,by="LAURING_ID",type = "left")

```


What would we expect?
Well here is our accuracy and the number of iSNV in each bin. A little back of the envelope calculation where we assume all variants between 2-5% have sensivity of 2% and all between 5-10% have sensitivity of 5%.

```{r,eval=F}
accuracy<-read.csv("../data/reference/accuracy_stringent.csv",stringsAsFactors = F)
trans_freq<-mutate(trans_freq,gc_ul1=meta$gc_ul[match(SPECID1,meta$SPECID)],gc_ul2=meta$gc_ul[match(SPECID2,meta$SPECID)])
gc<-10^c(3:5)
freq<-c(0,0.02,0.05,0.01)
trans_freq<-plyr::adply(trans_freq,1,function(x){
  x$gc_ul= max( gc[ which(gc<x$gc_ul.1) ])
  x$freq = max(freq[which(freq<=x$freq1)])
  return(x)})


counts<-plyr::ddply(trans_freq,~gc_ul+freq,summarize,count=length(gc_ul))
counts<-subset(counts,freq %in% c(0.02,0.05))

expected<-plyr::join(accuracy,counts)

expected<-mutate(expected,missed=round((count/sensitivity)-count))

five <- sum(expected$missed[expected$freq==0.05])
two <- sum(expected$missed[expected$freq==0.02])

0.02*two+0.05*five
```

We would expect 9. We find 17, but clearly NS is an outlier.


Now we will restrick our analysis only to sites that are polymporphic in the donor.

```{r}
trans_freq.comp<-read_csv("../data/processed/secondary/transmission_pairs_freq.poly.donor.csv")

trans_freq %>%  .$pair_id %>% unique()%>% length()->pair_id_whole

trans_freq.comp   %>% .$pair_id %>% unique()%>% length()->pair_id_poly

write_to_summary("No donor iSNV:",pair_id_whole-pair_id_poly)

write_to_summary("Final pair count:",pair_id_poly)
```

# Transmitted minority variants

```{r}
trans_freq.comp %>% filter(freq1<0.5)->dminor

dminor$chr<-factor(dminor$chr,levels = rev(c("PB2","PB1","PA","HA","NP","NR","M","NS"))) # Set the segments as factors with PB2 on top
chrs<-read.csv("../data/reference/segs.csv",stringsAsFactors = F) # get the start and stop of each OR for each segment (2014-2015 used as reference)

chrs$chr<-factor(chrs$chr,levels=levels(dminor$chr)) # set factors on the is meta data

#Add donor class
dminor<-left_join(dminor,select(qual,SPECID,mutation,class_factor),by=c("SPECID1" = "SPECID","mutation"))


genome_loc.p<-ggplot(dminor,aes(x=pos,y=as.factor(chr)))+
  geom_point(size=5,alpha=0.5,aes(color=class_factor,shape=found))+
  geom_segment(data=chrs,aes(x = start, y = chr, xend = stop, yend = chr))+
  ylab("")+xlab("")+
  scale_color_manual(name="",values=cbPalette[c(1,3)])+ 
  theme(axis.ticks =element_blank(), 
        axis.line.x = element_blank(),axis.line.y=element_blank())+
  scale_x_continuous(breaks=c())
genome_loc.p+ggtitle("transmitted minor")

dminor<-mutate(dminor,Trans = ifelse(found==T,"Transmitted","Not_Transmitted"))
dminor_table<-dminor %>% count(class_factor,Trans) %>%
  group_by(Trans) %>%
  spread(Trans,n) %>% rowwise()%>%
  mutate(Total = sum(Transmitted,Not_Transmitted))

dminor_table %>% select(ends_with("Transmitted")) %>% as.matrix() %>% fisher.test(.)
kable(dminor_table)

```



### Probability of transmission as a function of donor frequency

```{r}
trans_plot<-function(variant.df){
  
  logit=glm(formula =found~freq1,family=binomial(logit),data=variant.df )

  variant.df$prob=logit$fitted.values


  fit=ggplot(variant.df,aes(x=freq1,y=as.numeric(found)))+geom_point(alpha=0.1)+geom_line(aes(x=freq1,y=prob))+xlab("")+ylab("Probability")+theme(legend.position=c(0.9,0.5))#+geom_point(data =trans_freq.comp,aes(x=freq1,y=as.numeric(found)+0.02),color="red",alpha=0.1)+geom_line(data=trans_freq.comp,aes(x=freq1,y=prob,color='red'))

  found=ggplot(subset(variant.df,found==T),aes(x=freq1))+geom_histogram(position='dodge',binwidth = 0.02,color="white")+scale_y_log10()+xlab("")+theme(legend.position='none')+ylab("")
  
  lost=ggplot(subset(variant.df,found==F),aes(x=freq1))+geom_histogram(position='dodge',binwidth = 0.02,color="white")+scale_y_log10()+xlab("Frequency in first sample")+theme(legend.position='none')+ylab("")



grid.newpage()
  print(found, vp=viewport(0.9, 0.2, x=0.45, y=0.8))
  print(fit+guides(fill=FALSE), vp=viewport(width = 0.9,height =0.6, x=0.45, y=0.45))
  print(lost, vp=viewport(0.9, 0.2, x=0.45, y=0.1))
}

```

```{r}
trans_plot(trans_freq.comp)
```
### No frequency cut off
```{r}
no_cut.trans_freq<-read_csv("../data/processed/secondary/no_cut_trans_freq.csv",
                 col_types = list(
                 ENROLLID1= col_character(),
                 ENROLLID2= col_character(),
                 SPECID1 = col_character(),
                 SPECID2 = col_character(),
                 pair_id = col_double()))
```

```{r}
no_cut.trans_freq.p<-ggplot(no_cut.trans_freq,aes(x=freq1,y=freq2))+geom_point()+xlab("Frequency in donor")+ylab("Frequency in recipient")

no_cut.trans_freq.p
no_cut.trans_freq.p.log<-no_cut.trans_freq.p +
  scale_y_log10(breaks=c(0.0001,0.001,0.01,0.1,1))+
  scale_x_log10(breaks=c(0.0001,0.001,0.01,0.1,1))
no_cut.trans_freq.p.log

# save_plot("./Figures/Supplemental_Figure7A.pdf", no_cut.trans_freq.p,
#           base_aspect_ratio = 1.3)
# embed_fonts("./Figures/Supplemental_Figure7A.pdf")

# save_plot("./Figures/Supplemental_Figure7C.pdf", no_cut.trans_freq.p.log,
#           base_aspect_ratio = 1.3)
# embed_fonts("./Figures/Supplemental_Figure7C.pdf")

write.csv(x = select(no_cut.trans_freq,SPECID1,SPECID2,freq1,freq2),"./Figures/data/Supplemental_Figure7C.csv")


```


```{r}
no_cut.trans_freq.comp<-read_csv(
  "../data/processed/secondary/no_cut_transmission_pairs_freq.poly.donor.csv",
  col_types = list(
                 ENROLLID1= col_character(),
                 ENROLLID2= col_character(),
                 SPECID1 = col_character(),
                 SPECID2 = col_character(),
                 pair_id = col_double()))

```

```{r}
trans_plot(no_cut.trans_freq.comp)


```
```{r}
add_prob<-function(variant.df){
logit=glm(formula =found~freq1,family=binomial(logit),data=variant.df )

variant.df$prob=logit$fitted.values
return(variant.df)
}

trans_freq.comp<-add_prob(trans_freq.comp)
no_cut.trans_freq.comp=add_prob(no_cut.trans_freq.comp)

fit=ggplot()+geom_line(data=trans_freq.comp,aes(x=freq1,y=prob),color="black")+geom_line(data=no_cut.trans_freq.comp,aes(x=freq1,y=prob),color=cbPalette[5])+xlab("Frequency in Donor")+ylab("Probability")
fit
```

### Community pairs


```{r}
community_pairs.freq<-read_csv("../data/processed/secondary/community_pairs.freq.csv",
                 col_types = list(
                 ENROLLID1= col_character(),
                 ENROLLID2= col_character(),
                 SPECID1 = col_character(),
                 SPECID2 = col_character()
               ))
community_pairs.freq.comp<-read_csv(
  "../data/processed/secondary/community_pairs_freq.poly.donor.csv",
  col_types = list(
                 ENROLLID1= col_character(),
                 ENROLLID2= col_character(),
                 SPECID1 = col_character(),
                 SPECID2 = col_character(),
                 freq2 = col_double()
               ))


```


The community plot is flatter than the household one but - the log scale makes the histograms a bit miss leading.

```{r}
trans_plot(community_pairs.freq.comp)
ggplot(community_pairs.freq.comp,aes(x=freq1,fill=found))+geom_histogram(color='white',position="dodge")+scale_fill_manual(values=cbPalette[c(2,4)])+xlab("Frequency in donor")
```


```{r}
community_pairs.freq.comp=add_prob(community_pairs.freq.comp)

fit=ggplot()+geom_line(data=trans_freq.comp,aes(x=freq1,y=prob),color="black")+geom_line(data=community_pairs.freq.comp,aes(x=freq1,y=prob),color=cbPalette[5])+xlab("Frequency in Donor")+ylab("Probability")
fit
```

To get an idea of the spread around this line will sample the community pairs 1000 times each time grabbing the same number of transmission pairs as in the household data.

```{r}
p<-length(unique(trans_freq.comp$pair_id))
com_pairs.sampled<-sample_trans(community_pairs.freq.comp,1000,p)
```




```{r}
community.area_all<-plyr::ddply(com_pairs.sampled,~freq1,summarize,low.95=quantile(prob,na.rm=T,probs=0.025),high.95=quantile(prob,na.rm=T,probs=0.975),low.50=quantile(prob,na.rm=T,probs=0.25),high.50=quantile(prob,na.rm=T,probs=0.75))

trans.com.plot.sampled<-ggplot()+
  geom_ribbon(data=community.area_all,aes(x=freq1,ymin=low.95,ymax=high.95),alpha=0.6,fill=cbPalette[1])+
  geom_ribbon(data=community.area_all,aes(x=freq1,ymin=low.50,ymax=high.50),alpha=0.9,fill=cbPalette[1])+
  xlab("")+ylab("Probability transmitted")+
  geom_line(data=trans_freq.comp,aes(x=freq1,y=prob))+xlab("Frequency in Donor")

write.csv(select(community.area_all,frequency.in.donor=freq1,top.95.percentile=high.95,
                 bottom.95.percentile=low.95,
                 top.50.percentile=high.50,
                 bottom.50.percentile=low.50),"./Figures/data/Figure3E.area.csv")
write.csv(select(trans_freq.comp,frequency.in.donor = freq1,probability.of.transmission=prob,transmitted=found),
          "./Figures/data/Figure3DEF.line.csv")

trans.com.plot.sampled

save_plot("./Figures/Figure3E.pdf", trans.com.plot.sampled,
          base_aspect_ratio = 1.3)
embed_fonts("./Figures/Figure3E.pdf")
```


```{r}
#save(trans_freq.p,trans.com.plot,trans.com.plot.sampled,file="./transmission_setup_plots.RData")
```


# Appendex 

What if we don't require the community pairs to have a time dependency. The recipient could be sick first.

```{r,eval=F}
possible_pairs<-read.csv("./every_possible_pair.csv",stringsAsFactors = F) # All possible pairs (1 SPECID/person)
possible_pairs %>% mutate( Household1 = meta$HOUSE_ID[match(x = SPECID1,meta$SPECID)],
                            Household2 = meta$HOUSE_ID[match(x = SPECID2,meta$SPECID)],
                           Household = Household1==Household2) -> possible_pairs

all_community_pairs<-subset(possible_pairs,Household==F) # not household pairs
all_community_pairs$pair_id=1:nrow(all_community_pairs)
```


```{r,eval=F}
all_community_pairs.trans.donor<-subset(all_community_pairs,ENROLLID1 %in% trans_freq.comp$ENROLLID1)
community_pairs.trans.donor<-mutate(all_community_pairs.trans.donor,SPECID1=useful_trans_pairs$SPECID1[match(ENROLLID1,useful_trans_pairs$ENROLLID1)]) # make sure we are using the same SPECID

all_community_pairs.trans.donor.freq<-plyr::adply(all_community_pairs.trans.donor,1,
                                              function(x) get_freqs(c(x$SPECID1,x$SPECID2),qual),
                                              .parallel = T) 


all_community_pairs.trans.donor.freq.comp<-polish_freq(all_community_pairs.trans.donor.freq,freq1,0.02) # only polymorphic sites in sample 1 
all_community_pairs.trans.donor.freq.comp$found=all_community_pairs.trans.donor.freq.comp$freq2>0.02 # was it found in the second sample
```

```{r,eval=F}
all_com_pairs.trans.donor.sampled<-com_sample_trans(all_community_pairs.trans.donor.freq.comp,1000,trans_pair_specid$SPECID1)
```

```{r,eval=F}
all_community.area<-plyr::ddply(all_com_pairs.trans.donor.sampled,~freq1,summarize,low.95=quantile(prob,na.rm=T,probs=0.025),high.95=quantile(prob,na.rm=T,probs=0.975),low.50=quantile(prob,na.rm=T,probs=0.25),high.50=quantile(prob,na.rm=T,probs=0.75))

trans.all_com.plot<-ggplot()+geom_ribbon(data=all_community.area,aes(x=freq1,ymin=low.95,ymax=high.95),alpha=0.6,fill=cbPalette[1])+geom_ribbon(data=all_community.area,aes(x=freq1,ymin=low.50,ymax=high.50),alpha=0.9,fill=cbPalette[1])+xlab("")+ylab("Probability transmitted")+geom_line(data=trans_freq.comp,aes(x=freq1,y=prob))+xlab("Frequency in Donor")
trans.all_com.plot
```
