---
title: "frequency distribution"
output: html_document
---

What is spread of our observed frequency given the known frequency. To do this I will be using the obseverd frequencies from our benchmarking experiment. 

```{r}
freq<-read.csv("~/Desktop/benchmarking_frequency_copy.csv",stringsAsFactors = F,skip = 1)

require(plyr)

freq<-mutate(freq,spread = freq.var - exp.freq)
require(ggplot2)
freq<-subset(freq,exp.freq>=0.02)
ggplot(freq,aes(x=spread))+geom_histogram(color="white",binwidth = 0.01)
```


Now what if we normalize so that the first dilution titer 5 exp.freq 5 is 5%. This removes any error due to the qpcr or pipetting

```{r}

t<-ddply(freq,~mutation,function(x){
  first = subset(x,exp.freq==0.05 & gc==5)
  n_factor=first$freq.var/0.05 # normalization factor
  x$n_factor=n_factor
  x<-mutate(x,freq.n = freq.var/n_factor,spread.n = freq.n-exp.freq)
  return(x)
})



ggplot(subset(t,gc+exp.freq!=5.05 ),aes(x=spread.n))+geom_histogram(color="white",binwidth = 0.01)#+facet_wrap(~gc)
```

As expected distribution is dependent on the titer of the sample
```{r}
ggplot(subset(t,gc+exp.freq!=5.05 ),aes(x=spread.n,fill=as.factor(gc)))+geom_histogram(color="white",position='dodge',binwidth = 0.01)
```


```{r}
ggplot(subset(t,gc+exp.freq!=5.05 ),aes(x = gc,y=abs(spread.n)))+geom_point()#+scale_x_reverse()
```

It seems the expected frequency also play a role. Probably because there is more room to roam from higher frequency variants. We are only plotting those we detect.
```{r}
ggplot(subset(t,gc+exp.freq!=5.05 ),aes(x = exp.freq,y=abs(spread.n)))+geom_point()#+scale_x_reverse()
```



# Fitting

These should be normal error. The distributions are skewed because we are looking at minor alleles and are bound on the lower end by a limit of detetion. I'll fit for the 10^4 data set using method of moments excluding the estreem point.Should the mean be 0. I'll round to 0 it was 0.002. 


```{r}
fit4<-subset(t,gc==4 & spread.n<0.1)
mean(fit4$spread.n)
v = var(fit4$spread.n)
x= seq(-0.1,0.1,0.01)
plot(x,dnorm(x,mean=0,sd = sqrt(v)))
```